---
title:  Accelerated Failure Time Models with ciTools
author: John Haman
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accelerated Failure Time Models with ciTools}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, message = FALSE }
library(dplyr)
library(ggplot2)
library(knitr)
library(ciTools)
library(here)
set.seed(20180925)
```

*Disclaimer:* `ciTools` makes three assumptions about your model:

1 - no missing data in the "newdata" dataframe, `df`.

2 - distribution is one of weibull, lognormal, loglogistic, or exponential.

3 - regression model is unweighted and without random effects.


The purpose of this vignette is to introduce and discuss new `ciTools`
capabilities for handling accelerated failure time (AFT) models. Some
of the new AFT methods in `ciTools` are more informative than methods
for previous models, and will inform future development decisions that
will be made in `ciTools`. In particular, `ciTools` now supports
intervals for estimated survival time probabilities and quantiles for
a range of common AFTs.

## The Accelerated Failure Time Model

The accelerated failure time model is, like a generalized linear
model (GLM), an extension of the standard linear model that
accounts for specific types of data and non-linearity. AFTs
constitute and important class of models as they can handle
censored, highly skewed data -- exactly the type of data one would
expect to collect when analyzing the failure times of a machine, or
the survival times of a group of patients under study.

AFTs are special in the field of survival/reliability analysis in
that they are fully parametric models. This provides power to do
certain inferences such as the estimation of tail probabilities
that would be difficult in a non or semi-parametric framework. The
trade-off made is that a specific distribution for survival times
must be assumed, and this assumption may be incorrect.

The structure of accelerated failure time models is as follows. We
observe a vector of survival times (failure times, in the
reliability literature) $T$ given a data matrix $X$. We assume the
$\log$ of the survival times is affected linearly by the
covariates of $X$. Because $T$ is non-negative, we model the effect
of the linear predictor $X\beta$ on $\log(T)$. The model is

$$
F(T|X) = F \left(
\frac{\log(T) - X\beta}{\sigma}
\right).
$$

$F$ denotes a vectorized, standard distribution function; $X\beta$
is called the linear predictor; and $\sigma$ is called the scale
parameter. $S(T|X)$ is called he survivor function; the probability
that a unit will fail after time $T$. $S(T|X) = 1 - F(T|X)$. Thus
the AFT model is a family of log-linear models. Examples of $F$
that are common are the standard normal, standard logistic, and
standard smallest extreme value distribution functions (Meeker and
Escobar, Ch. 4). We can more clearly write the model as

$$
\log(T) = X\beta + \sigma \varepsilon,
$$

where $\varepsilon \sim F$ to make the linear effect of $\beta$
on $\log(T)$ a bit more apparent.

Like generalized linear  models, survival models are  fit through a
maximum likelihood procedure. This is most useful in that it allows
a  practitioner  to  specify   censored  data  in  the  statistical
model.  That AFTs  are fully  parametric and  may account  for data
censoring were  primary reasons  for adding  them to  `ciTools`. We
assume that AFTs are fit in  **R** with the `survreg` function from
the `survival` library.

### Examples of AFTs

Four examples of AFT models are presented, which are covered
completely by `ciTools`. This list of AFT models is not exhaustive,
as other models are available. See the `flexsurv` package, for
example. Models in the `flexsurv` package do not presently receive
treatment by `ciTools`.

1. **Lognormal**: Let $\varepsilon \sim N(0,1)$. Then $\log(T) =
X\beta + \sigma \varepsilon$ and $T$ is said to be lognormal with
parameters $X\beta$ and $\sigma$. Confidence intervals for the
following parameters are available in `ciTools`.

$$
E(T|X) = \exp(X \beta + \frac{\sigma^2}{2}) \qquad (\text{expected time to failure})
$$

$$
\text{median}(T|X) = \exp(X\beta)  \qquad (\text{median time to failure})
$$

$$
S(T|X) = 1 - \Phi \left(
\frac{\log(T) - X\beta}{\sigma}
\right) \qquad \Phi \sim \text{std. Normal CDF}
$$

$$
F^{-1}_p(T|X) = \exp(X\beta + \Phi^{-1}(p) \sigma) \qquad (\text{level p quantile of failure time distribution})
$$

2. **Weibull**: Let $\varepsilon$ possess smallest extreme value
distribution. We write $\varepsilon \sim SEV$ with
$F_{SEV}(\varepsilon) = 1-\exp(-\exp(\varepsilon))$. If $\log(T) =
X\beta + \sigma\varepsilon$, then $T$ is *weibull* distributed with
scale parameter $\sigma$ and location parameter $\exp(X\beta)$ in
the location-scale parameterization used in the `survival`
package. This parameterization differs from the one used in
`{p/d/q/r}weibull`, see `help(survreg)` for details.

$$
E(T|X) = \exp(X\beta)\Gamma(1 + \sigma)
$$

$$
\text{median}(T|X) = \exp(X\beta + F^{-1}_{SEV}(0.5) \sigma)
= \exp(X\beta)(\log(2))^{\sigma}
$$

$$
F^{-1}_p(T|X) = \exp(X\beta + F^{-1}_{SEV}(p) \sigma)
= \exp(X\beta)(-\log(1-p))^{\sigma}
$$

$$
S(T|X) = \exp(-\exp(z)), \qquad  z = \frac{\log(T) - X\beta}{\sigma}
$$

3. **Exponential**: Like the weibull distribution, except scale
parameter $\sigma$ is fixed to $1$.

$$
E(T|X) = \exp(X\beta)
$$

$$
\text{median}(T|X) = \exp(X\beta + F^{-1}_{SEV}(0.5))
= \exp(X\beta)\log(2)
$$

$$
F^{-1}_p(T|X) = \exp(X\beta + F^{-1}_{SEV}(p))
= \exp(X\beta)(-\log(1-p))
$$

$$
S(T|X) = \exp(-\exp(z)), \qquad  z = \log(T) - X\beta
$$

4. **Loglogistic**: Let $\varepsilon \sim \text{Logistic}$. That is,
$F(\varepsilon) = \frac{\exp(\varepsilon)}{1 + \exp(\varepsilon)}$,
the standard logistic distribution. Then $\log(T) = X\beta +
\sigma\varepsilon$, and $T$ is *loglogistic* distributed with scale
parameter $\sigma$ and location parameter $X\beta$.

$$
E(T|X) = \exp(X\beta)\Gamma(1 + \sigma)\Gamma(1 - \sigma)
$$

$$
\text{median}(T|X) = \exp(X\beta)
$$

$$
F^{-1}_p(T|X) = \exp(X\beta + \sigma F^{-1}_{Logistic}(p))
$$

$$
S(T|X) = 1 - F_{Logistic} \left(
\frac{\log(T) - X\beta}{\sigma}
\right)
$$

Note that the median of  each conditional failure time distribution
is   technically  the   level   $p=0.5$  quantile   of  that   same
distribution. For this reason, confidence intervals for medians are
calculated with `add_quantile()`.

## AFT Uncertainty Intervals

In the analysis of AFT models, statisticians have several options for
making predictions. `predict.survreg` for example, allows one to
predict the median failure time, or any other quantile. These
predictions have the same units as the original time scale (time to
death or failure). Additionally, `predict.survreg` can also output the
corresponding value of the linear predictor for a given point in the
factor space.

`ciTools` hopes to clarify survival times prediction for AFT models
by relegating  prediction of the  expected (mean) survival  time to
`add_ci.survreg`, and prediction of the median (or any quantile) of
the  survival  time  distribution to  `add_quantile.survreg`.  Thus
`add_ci.survreg` is in line with other `add_ci` S3 methods provided
by  `ciTools`  by  only  providing  confidence  intervals  for  the
expected response conditioned on the predictors.

There are three popular methods  for forming confidence intervals in
this case: parametrically, using either the (1) delta method or (2)
likelihood ratios, or (3) through a bootstrap resampling procedure.
In `ciTools`,  we generally favor parametric  methods, except where
it makes sense  to include bootstrap methods as options,  as is the
case with  many mixed effects  models, where bootstrap  methods are
seen as less controversial than parametric interval methods.

We have studied these three  techniques for interval estimation, and
found that  the delta method  offers the best combination  of speed
and accuracy for users. Therefore, the  delta method is used as the
basis of  all interval estimation  procedures in `ciTools`  for AFT
models.  This  is at  odds with the  recommendations of  Meeker and
Escobar  in  favor  of   likelihood  based  intervals,  however  we
implement delta method  based intervals as they are  much easier to
write  for  multivariate   models  and  do  not   suffer  from  any
convergence issues.  Compared to bootstrap intervals,  delta method
intervals are faster and have  similar probability coverage in many
scenarios.

## Example.

Data are collected  on the failure times of a  new spring installed
in a car. The spring can be mounted in two types of cars, an SUV or
a  sedan.  An  additional  variable, ambient  temperature was  also
recorded. Experimental  vehicles were fitted with  the new springs,
and the  vehicles are placed  into an observational  study. Vehicles
with the  new springs  were driven,  and the  failure times  of the
springs were noted until the conclusion  of the test time. The test
concluded after $2000$ cumulative hours  of testing.  At this time,
all surviving springs at marked  as right censored at $t=2000$. All
data are notional.


```{r include = FALSE, cache = TRUE}
library(SPREDA)
car <- rep(c("suv", "sedan"), 25)
temp <- seq(40,100, length.out = 50)
time <- exp(0.33 + 0.2 * ifelse(car == "sedan", 0, 1) + 0.08 * temp + rsev(50))
time <- ifelse(time < 2000, time, 2000)
failure <- ifelse(time < 2000, 1, 0)
dat <- data.frame(temp = temp,
                  car = car,
                  time = time,
                  failure = failure)
```

The time variable indicates of the  number of hours driven before a
spring failure is  observed. If *failure = 1*, a  spring failure is
observed at the indicated time.

```{r cache = TRUE}
kable(head(dat))
```

```{r cache = TRUE, fig.width = 5, fig.height= 5, warning = FALSE}
ggplot(dat, aes(x = temp, y = time)) +
    geom_point(aes(color = factor(failure)))+
    ggtitle("Censored obs. in red") +
    theme_bw()
```

Seven of the observations are censored at $t = 2000$. This means we
assume those $7$ springs would eventually fail at some later point
in time had we chosen to run the study for a longer period of
time. We fit a weibull model to the data. By default, `Surv` will
infer (correctly, in this case) that our observations are right
censored at $t=2000$. Check the documentation of `Surv` for how to
coordinate a different censoring regime -- `survreg` is very
flexible in the types of censoring allowed (another advantage of AFT
models). Other distributions (exponential, lognormal, loglogistic)
are available in `survreg` for parametric analysis, and receive
treatment in `ciTools`, but we stick the weibull model for the
example.

```{r cache = TRUE}
(fit <- survreg(Surv(time, failure) ~ temp + car, data = dat)) ## weibull dist is default
```

The output  of `survreg` indicates that  the model on the  whole is
significantly  better   than  one   which  does  not   include  any
covariates.  Maximum  likelihood   estimates  of  coefficients  are
displayed as  well. We can  analyze the model graphically  with the
help of `ciTools`. The `summary` function can be called on `fit` to
show some  additional information about the  model coefficients. We
calculate confidence  and prediction intervals, and  append them to
the original data set.

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
with_ints <- ciTools::add_ci(dat,fit, names = c("lcb", "ucb")) %>%
    ciTools::add_pi(fit, names = c("lpb", "upb"))
```

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
kable(head(with_ints))
```

The output of `ciTools`'s functions is always the inputted data with
the requested statistics attached. The inputted data can be original
data or a data frame of new observations. For this model fit, `add_ci`
calculates conditional means (denoted `mean_pred` in the data frame)
and `add_pi` calculates conditional medians (`median_pred` in the data
frame).

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
ggplot(with_ints, aes(x = temp, y = time)) +
    geom_point(aes(color = car)) +
    facet_wrap(~car)+
    theme_bw() +
    ggtitle("Model fit with 95% CIs and PIs",
            "solid line = mean, dotted line = median") +
    geom_line(aes(y = mean_pred), linetype = 1) +
    geom_line(aes(y = median_pred), linetype = 2) +
    geom_ribbon(aes(ymin = lcb, ymax = ucb), alpha = 0.5) +
    geom_ribbon(aes(ymin = lpb, ymax = upb), alpha = 0.1)
```

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
probs <- ciTools::add_probs(dat, fit, q = 500,
                            name = c("prob", "lcb", "ucb"),
                            comparison = ">")
```

We can calculate the estimated survival probabilities as
well. Below, we calculate the probability of a spring failing after
$t = 500$ (alternatively, the probability of a spring not failing
before $t = 500$). This is not a new feature to `ciTools`, what's
special to `survreg` methods is that `ciTools` will additionally
compute confidence intervals for the estimated conditional survival
probabilities.


```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
ggplot(probs, aes(x = temp, y = prob)) +
    ggtitle("Estimated prob. of avg. spring lasting longer than 500 hrs.") +
    ylim(c(0,1)) +
    facet_wrap(~car)+
    theme_bw() +
    geom_line(aes(y = prob)) +
    geom_ribbon(aes(ymin = lcb, ymax = ucb), alpha = 0.5)
```

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
quants <- ciTools::add_quantile(dat, fit, p = 0.90,
                                name = c("quant", "lcb", "ucb"))
```

Furthermore, we can calculate quantiles of the distribution of
failure times given the covariates in `dat`. Again, the special
sauce for AFT models comes when `ciTools` also tacks on confidence
intervals for the estimated quantiles. Here, we show the estimated
$0.90$ quantile, conditioned on the covariate information, with
confidence intervals. One may use `add_quantile` to calculate the
median failure time, or any other quantile.

```{r cache = TRUE, fig.width = 8, fig.height = 5, warning = FALSE}
ggplot(quants, aes(x = temp, y = time)) +
    geom_point(aes(color = car)) +
    ggtitle("Estimated 90th percentile of condtional failure distribution, with CI") +
    facet_wrap(~car)+
    theme_bw() +
    geom_line(aes(y = quant)) +
    geom_ribbon(aes(ymin = lcb, ymax = ucb), alpha = 0.5)
```

## The Delta Method for Regression Models

Here are the mathematical details for calculating the above confidence
intervals. For AFT models, we calculate confidence intervals with the
delta method (Prediction intervals are calculated using different
methods, discussed later).

Let $\boldsymbol{\theta} = (\beta_0,
\beta_1, \ldots, \beta_p, \sigma)$ be the vector of maximum likelihood
parameter estimates for the statistical model. We wish to form
confidence intervals for continuous and twice differentiable functions
of $\boldsymbol{\theta}$, say
$\mathbf{g}(\boldsymbol{\theta})$. Because
$\hat{\boldsymbol{\theta}}_{ML}$ is a maximum likelihood estimator of
$\boldsymbol{\theta}$, $\mathbf{g}(\hat{\boldsymbol{\theta}_{ML}})$ is
a maximum likelihood estimator of
$\mathbf{g}(\boldsymbol{\theta})$. In large samples,
$\mathbf{g}(\hat{\boldsymbol{\theta}})$ is approximately Normally
distributed with mean $\mathbf{g}(\boldsymbol{\theta})$ and
variance-covariance matrix

$$
\Sigma_{\hat{g}} = \left[ \frac{\partial \mathbf{g}
(\boldsymbol{\theta})}{\partial \boldsymbol{\theta}}\right]^T
\Sigma_{\hat{\boldsymbol{\theta}}} \left[ \frac{\partial \mathbf{g}
(\boldsymbol{\theta})}{\partial \boldsymbol{\theta}}\right].
$$

This approximation is based on the assumption that
$\mathbf{g}({\hat{\boldsymbol{\theta}}})$ is linear in
$\hat{\boldsymbol{\theta}}$ in a region near
$\boldsymbol{\theta}$. The larger the sample, the better, because
the variation in $\hat{\boldsymbol{\theta}}$ decreases with sample
size and thus the region over which $\hat{\boldsymbol{\theta}}$
varies is correspondingly smaller. If the region is small enough,
the approximation is adequate.

Mathematically, the delta method is a statistical rebranding of the
a Taylor series expansion for
$\mathrm{Var}[\mathbf{g}(\hat{\boldsymbol{\theta}})]$ . We will use
the delta method to form confidence intervals for functions of
$\boldsymbol{\theta}$: expected values, quantiles, and survivor
functions.

## Expected Values

Because it is somewhat easier to explain confidence intervals for the
mean if we have a particular model in mind, suppose for the moment
that we fit a *Weibull* AFT model. The expected survival time,
$\mathrm{E}[T|X]$, written as a function of $\boldsymbol{\theta}$, is
$\mathbf{g}(\boldsymbol{\theta}) = \exp(X\beta)\Gamma(1 + \sigma)$. We
form a confidence interval for this mean survival time based on the
estimated regression coefficients ($\hat{\boldsymbol{\beta}}$, and
$\hat{\sigma}$) from `survreg`.  Due to some quirks in numerical
optimization, it is often advantageous to reparameterize the scale
parameter as $\delta = \log(\sigma)$ in the model.  Let $x$ denote a
point in the factor space at which we wish to calculate the expected
failure time. The relevant derivatives are

$$
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial \boldsymbol{\beta}} =
\exp(x^T \boldsymbol{\beta}) x^T \Gamma(1 + \exp(\delta)) ,\qquad i = 0, \ldots, p,
$$
and
$$
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial \delta} =
\exp(x^T \boldsymbol{\beta}) \Gamma(1 + \exp(\delta)) \psi(1 + \exp(\delta)) \exp(\delta),
$$

where $\psi(\cdot)$ denotes the digamma function. Let
$\frac{\partial\mathbf{g}(\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}} = \left(\frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_0}, \frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_1}, \frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_2}, \ldots,
\frac{\partial \mathbf{g}(\boldsymbol{\theta})}{\partial \beta_p},
\frac{\partial \mathbf{g}(\boldsymbol{\theta})}{\partial
\delta}\right)^T$.

The standard error of the expected survival time estimate is

$$
\mathrm{s.e.}(\mathbf{g} (\hat{\boldsymbol{\theta}})) = \sqrt{\left[
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}}\right]^T \Sigma_{\hat{\boldsymbol{\theta}}}
\left[ \frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}}\right]}.
$$

An approximate $100\times(1 - \alpha)\%$ confidence interval for
$\mathbf{g}(\boldsymbol{\theta}) = \mathrm{E}[T]$
based on the large sample standard Normal approximation of
$Z_{\log(\mathbf{g}(\hat{\boldsymbol{\theta}}))} =
\frac{\log(\mathbf{g}(\hat{\boldsymbol{\theta}})) -
\log(\mathbf{g}(\boldsymbol{\theta}))}{s.e.(\mathbf{g}
(\hat{\boldsymbol{\theta}}))}$ is

$$
\left[\mathrm{lower}, \mathrm{upper} \right] =
\left[\mathbf{g}(\hat{\boldsymbol{\theta}})/w,
\mathbf{g}(\hat{\boldsymbol{\theta}}) \times w \right],
$$

where $w = \exp(z_{1-\alpha/2} \times \mathrm{s.e.}(\mathbf{g}
(\hat{\boldsymbol{\theta}})) /
\mathbf{g}(\hat{\boldsymbol{\theta}}))$.

Confidence intervals for the expected response of other AFT models
may be calculated similarly, except that the function $\mathbf{g}$
depends on the response distribution. For other functions of
$\boldsymbol{\theta}$ such as the survivor function or response
quantile, we apply the delta method as well.

### Simulation Setup

A simulation was conducted to investigate the performance of
uncertainty intervals for AFT models. We varied sample size,
distribution, and the proportion of observations censored. In all
simulations, a simple AFT models with one predictor was used. A time
censoring mechanism was assumed and set at one of three levels: no
censoring, mild censoring (30% of observations censored) or moderate
censoring (50% of observations censored).

The number of simulations for each combination of distribution and
censoring was $10,000$ (for the confidence intervals) or $5000$
(for prediction intervals, survival probabilities, and quantiles).

### Simulation for expected value CIs

We produce graphs of the performance of the delta method for
confidence intervals on the expected survival time. Below we show the
observed coverage probability as sample size, distribution, and
censoring proportion vary. The nominal coverage probability is set at
90% for all simulations.

We observe acceptable performance from the delta method in this
case. Larger amounts of censoring (30% and 50%) produce mostly lower
coverage probabilities for all distributions except the exponential
distribution. The worst case scenario appears to be small sample
Lognormal fits with moderate censoring. In this case, there is a near
10% gap between the nominal and observed coverage probabilities.

```{r include=FALSE, cache= TRUE}
path <- "./survreg_data"
exp0 <- read.csv(paste0(path,"/","exponential_sim_expect_ci_censoredp00.csv"))
exp1 <- read.csv(paste0(path,"/","exponential_sim_expect_ci_censoredp03.csv"))
exp2 <- read.csv(paste0(path,"/","exponential_sim_expect_ci_censoredp05.csv"))
exp_dat <- rbind(exp0, exp1, exp2) %>%
    mutate(distr = "Exponential")
loglog0 <- read.csv(paste0(path,"/",
                           "loglogistic_sim_expect_ci_scale025_censoredp00.csv"))
loglog1 <- read.csv(paste0(path,"/",
                           "loglogistic_sim_expect_ci_scale025_censoredp03.csv"))
loglog2 <- read.csv(paste0(path,"/",
                           "loglogistic_sim_expect_ci_scale025_censoredp05.csv"))
loglog_dat <- rbind(loglog0, loglog1, loglog2) %>%
    mutate(distr = "Loglogistic") %>%
    dplyr::select(-c(scale))
lognorm0 <- read.csv(paste0(path,"/","lognormal_sim_expect_ci_scale2_censoredp00.csv"))
lognorm1 <- read.csv(paste0(path,"/","lognormal_sim_expect_ci_scale2_censoredp03.csv"))
lognorm2 <- read.csv(paste0(path,"/","lognormal_sim_expect_ci_scale2_censoredp05.csv"))
lognorm_dat <- rbind(lognorm0, lognorm1, lognorm2) %>%
    mutate(distr = "Lognormal") %>%
    dplyr::select(-c(scale))
weibull0 <- read.csv(paste0(path,"/","weibull_sim_expect_ci_scale2_censoredp00.csv"))
weibull1 <- read.csv(paste0(path,"/","weibull_sim_expect_ci_scale2_censoredp03.csv"))
weibull2 <- read.csv(paste0(path,"/","weibull_sim_expect_ci_scale2_censoredp05.csv"))
weibull_dat <- rbind(weibull0, weibull1, weibull2) %>%
    mutate(distr = "Weibull") %>%
    dplyr::select(-c(scale))
ci_dat <- rbind(exp_dat, loglog_dat, lognorm_dat, weibull_dat)
```

```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE}
ggplot(filter(ci_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
                      ymax = cov_prob + 2*cov_prob_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("C.I. Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities (zoomed in)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)

```
```{r fig.width=8, fig.height=8, echo=FALSE, cache= TRUE}
ggplot(filter(ci_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    ylim(0,1) +
    geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
                      ymax = cov_prob + 2*cov_prob_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("C.I. Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities (zoomed out)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

Interval widths generally shrink to zero, which is expected. In one
case, Lognormal fits with sample size 20 and 50% censoring, the
interval widths are too large. This is due to too many unconverged
maximum likelihood estimates.

```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE}
ggplot(filter(ci_dat), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_errorbar(aes(ymin=int_width - 2*int_width_se,
                      ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("C.I Simulation, by Distribution and Censoring",
            subtitle = "Interval Widths") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.0), colour="blue", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```


## Survivor Function

Calculating confidence intervals for estimated probabilities requires
a bit more care to ensure that the confidence bounds lie in the (0,1)
interval. Because the mathematics of the confidence intervals for the
survivor function depend less on the actual distribution, we won't
focus on the Weibull model, and will treat all AFT models at once. The
predicted probability of survival at time $T$, $\hat{S}(T|X)$, written
as a function of $\boldsymbol{\theta}$, is
$\mathbf{g}(\boldsymbol{\theta}) = 1 - \Phi \left( \frac{\log(T) -
X\boldsymbol{\beta}}{\sigma}\right)$. Similar to the previous example,
we form a confidence interval for this probability of survival based
on the estimated regression coefficients ($\hat{\boldsymbol{\beta}}$
and $\hat{\sigma}$) from `survreg`. The maximum likelihood estimator
of the survivor function is $\hat{S}(T|X)_{ML} = \Phi \left(
\frac{\log(T) - X\hat{\beta}}{\hat{\sigma}}\right)$. The relevant
derivatives are

$$
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial \beta} =
f \left(\frac{\log(T) - x^T \boldsymbol{\beta}}{\sigma}\right) \times
\left(\frac{-x^T}{\sigma}\right),\qquad i = 0, \ldots, n,
$$

and

$$
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial \delta} =
f \left(\frac{\log(T) - x^T \boldsymbol{\beta}}{\sigma}\right) \times
\left(\frac{x^T \boldsymbol{\beta} - \log(T)}{\sigma} \right),
$$

where $f(\cdot)$ denotes the probability density function
corresponding to $\Psi$, and $x_i$ is a new observation$. Let
$\frac{\partial\mathbf{g}(\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}} = \left(\frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_0}, \frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_1}, \frac{\partial
\mathbf{g}(\boldsymbol{\theta})}{\partial \beta_2}, \ldots,
\frac{\partial \mathbf{g}(\boldsymbol{\theta})}{\partial \beta_p},
\frac{\partial \mathbf{g}(\boldsymbol{\theta})}{\partial
\delta}\right)^T$.

Due to the delta method, the mathematical form of the standard
errors of the estimated survival probability is as it was in the
previous example:

$$
\mathrm{s.e.}(\mathbf{g} (\hat{\boldsymbol{\theta}})) = \sqrt{\left[
\frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}}\right]^T \Sigma_{\hat{\boldsymbol{\theta}}}
\left[ \frac{\partial \mathbf{g} (\boldsymbol{\theta})}{\partial
\boldsymbol{\theta}}\right]}
$$

The obvious confidence interval based on the statistic
$\frac{\hat{S} - S}{\hat{s.e.}(\hat{S})}$ could be potentially a
very poor fit approximation. The approximation could be poor due to
a small to moderate number of failures in the data (Meeker and
Escobar, p. 190) or because the bounds of the confidence interval
could far exceed the interval $[0,1]$. The chosen solution is to
apply a transformation $u(\cdot)$ such that $\frac{u(\hat{S}) -
u(S)}{\hat{s.e.}(\hat{u})}$ is closer in distribution to a standard
Normal distribution. A transformation that achieves this is the
logistic transform.

$$
u(\hat{S}) = \log \left( \frac{\hat{S}}{1 - \hat{S}} \right)
$$

First, we find a confidence interval for $u(\hat{S})$, then
transform the endpoints of the interval to the $[0,1]$ through the
inverse logistic function to find a confidence interval for
$\hat{S}$.

$$
\left[\mathrm{lower}, \mathrm{upper} \right] =
\left[
\frac{\hat{S}}{\hat{S} + (1 - \hat{S}) \times w},
\frac{\hat{S}}{\hat{S} + (1 - \hat{S})/w}
\right]
$$

where $w = \exp \left( \frac{z_{1 - \alpha/2} \hat{s.e.} (\hat{S}) }
{\hat{S}(1 - \hat{S})}\right)$.

### Simulations for Survivor Function

Plots below show the performance of the delta method for uncertainty
intervals of the survivor function. Again we compare the observed
coverage probability with the 90% nominal probability. In contrast to
intervals for the mean, censoring appears to produce overly
conservative intervals. This is particularly clear in the case of the
Weibull distribution.

```{r, include = FALSE, cache= TRUE, warning = FALSE}
expp0 <- read.csv(paste0(path,"/","exponential_sim_prob_ci_censoredp00.csv"))
expp1 <- read.csv(paste0(path,"/","exponential_sim_prob_ci_censoredp03.csv"))
expp2 <- read.csv(paste0(path,"/","exponential_sim_prob_ci_censoredp05.csv"))
expp_dat <- rbind(expp0, expp1, expp2) %>%
    mutate(distr = "Exponential")
loglogp0 <- read.csv(paste0(path,"/",
                            "loglogistic_sim_prob_ci_scale025_censoredp00.csv"))
loglogp1 <- read.csv(paste0(path,"/",
                            "loglogistic_sim_prob_ci_scale025_censoredp03.csv"))
loglogp2 <- read.csv(paste0(path,"/",
                            "loglogistic_sim_prob_ci_scale025_censoredp05.csv"))
loglogp_dat <- rbind(loglogp0, loglogp1, loglogp2) %>%
    mutate(distr = "Loglogisic") %>%
    dplyr::select(-c(scale))
lognormp0 <- read.csv(paste0(path,"/","lognormal_sim_prob_ci_scale2_censoredp00.csv"))
lognormp1 <- read.csv(paste0(path,"/","lognormal_sim_prob_ci_scale2_censoredp03.csv"))
lognormp2 <- read.csv(paste0(path,"/","lognormal_sim_prob_ci_scale2_censoredp05.csv"))
lognormp_dat <- rbind(lognormp0, lognormp1, lognormp2) %>%
    mutate(distr = "Lognormal") %>%
    dplyr::select(-c(scale))
weibullp0 <- read.csv(paste0(path,"/","weibull_sim_prob_ci_scale2_censoredp00.csv"))
weibullp1 <- read.csv(paste0(path,"/","weibull_sim_prob_ci_scale2_censoredp03.csv"))
weibullp2 <- read.csv(paste0(path,"/","weibull_sim_prob_ci_scale2_censoredp05.csv"))
weibullp_dat <- rbind(weibullp0, weibullp1, weibullp2) %>%
    mutate(distr = "Weibull") %>%
    dplyr::select(-c(scale))
prob_dat <- rbind(expp_dat, loglogp_dat, lognormp_dat, weibullp_dat)
```

```{r, fig.width=8, fig.height=8, echo = FALSE, cache= TRUE, warning = FALSE}
ggplot(filter(prob_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
                      ymax = cov_prob + 2*cov_prob_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("C.I. for Survival Probability Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities (zoomed in)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

However on the (0,1) probability scale, we find acceptable performance.

```{r, fig.width=8, fig.height=8, echo = FALSE, cache= TRUE, warning = FALSE}
ggplot(filter(prob_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    ylim(0,1) +
    geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
                      ymax = cov_prob + 2*cov_prob_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("Survival Probability Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities (zoomed out)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

Intervals widths go to zero as sample size increases.

```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE, warning = FALSE}
ggplot(filter(prob_dat), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_errorbar(aes(ymin=int_width - 2*int_width_se,
                      ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("Survival Probability Simulation, by Distribution and Censoring",
            subtitle = "Interval Widths") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.0), colour="blue", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

## Prediction Intervals

We have not yet discussed prediction intervals, but `ciTools` also has
methods for creating two different types of prediction intervals. The
first type is the "naive" method of Meeker and Escobar. The naive
method simply forms a prediction interval based on $\alpha/2$ and
$1-\alpha/2$ quantiles of the estimated conditional distribution. The
method is naive in the sense that it does not account for uncertainty
in the estimates of the parameters $\boldsymbol{\beta}$ and
$\sigma$. This method is simple and works reasonably well in the
absence of censoring, as displayed in the plots below.

A slightly better method is included in `ciTools`, which we call the
simulation method. We generate prediction intervals for the next
failure time by a parametric bootstrap. This simulation method assumes
a multivariate normal distribution for the model coefficients
(excluding the scale parameter) and generates new responses that
account for this uncertainty in the model coefficients,
$\boldsymbol{\beta}$. This should make the bootstrap method slightly
better than the naive method in practice, though a bit more
computationally expensive. This is essentially what we have done to
produce prediction intervals for GLMs as well, just applied to the new
class of AFT models.

From the plot below, it's pretty clear that the bootstrap method
outperforms the naive method. The difference is most stark when there
is a moderate amount of censoring.

Further improvements beyond the naive and simulations methods may be
made. These techniques are detailed in Meeker and Escobar (Chapter
12), but have yet to be implemented in `ciTools`.


```{r include=FALSE, cache= TRUE}
exppi0 <- read.csv(paste0(path,"/","exponential_sim_pi_censoredp00.csv"))
exppi1 <- read.csv(paste0(path,"/","exponential_sim_pi_censoredp03.csv"))
exppi2 <- read.csv(paste0(path,"/","exponential_sim_pi_censoredp05.csv"))
exppi_dat <- rbind(exppi0, exppi1, exppi2) %>%
    mutate(distr = "Exponential")
loglogpi0 <- read.csv(paste0(path,"/",
                             "loglogistic_sim_pi_scale025_censoredp00.csv"))
loglogpi1 <- read.csv(paste0(path,"/",
                             "loglogistic_sim_pi_scale025_censoredp03.csv"))
loglogpi2 <- read.csv(paste0(path,"/",
                             "loglogistic_sim_pi_scale025_censoredp05.csv"))
loglogpi_dat <- rbind(loglogpi0, loglogpi1, loglogpi2) %>%
    mutate(distr = "Loglogistic") %>%
    dplyr::select(-c(scale))
lognormpi0 <- read.csv(paste0(path,"/","lognormal_sim_pi_scale2_censoredp00.csv"))
lognormpi1 <- read.csv(paste0(path,"/","lognormal_sim_pi_scale2_censoredp03.csv"))
lognormpi2 <- read.csv(paste0(path,"/","lognormal_sim_pi_scale2_censoredp05.csv"))
lognormpi_dat <- rbind(lognormpi0, lognormpi1, lognormpi2) %>%
    mutate(distr = "Lognormal") %>%
    dplyr::select(-c(scale))
weibullpi0 <- read.csv(paste0(path,"/","weibull_sim_pi_scale2_censoredp00.csv"))
weibullpi1 <- read.csv(paste0(path,"/","weibull_sim_pi_scale2_censoredp03.csv"))
weibullpi2 <- read.csv(paste0(path,"/","weibull_sim_pi_scale2_censoredp05.csv"))
weibullpi_dat <- rbind(weibullpi0, weibullpi1, weibullpi2) %>%
    mutate(distr = "Weibull") %>%
    dplyr::select(-c(scale))
pi_dat <- rbind(exppi_dat, loglogpi_dat, lognormpi_dat, weibullpi_dat)
```
```{r fig.width=8, fig.height=8, echo=FALSE, cache= TRUE}
ggplot(filter(pi_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_line(aes(y = cov_prob_boot), size = 1.5, color = "blue") +
    ## geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
    ##                   ymax = cov_prob + 2*cov_prob_se), width=.1) +
    ## geom_errorbar(aes(ymin=cov_prob_boot - 2*cov_prob_boot_se,
    ##                   ymax = cov_prob_boot + 2*cov_prob_boot_se),
    ##               width=.1, color = "blue") +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("P.I. Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities. Naive method (black). Simulation method (blue) ") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE}
ggplot(filter(pi_dat), aes(x=sample_size, y=cov_prob)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    ylim(0,1) +
    geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se,
                      ymax = cov_prob + 2*cov_prob_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Coverage Probs +/- 2 SEs") +
    ggtitle("P.I. Simulation, by Distribution and Censoring",
            subtitle = "Coverage Probabilities (zoomed out)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) +
    facet_grid(distr~censoredp)
```

Unlike confidence intervals, interval widths of prediction intervals
should not shrink to zero as sample size is increased. Instead we
should observe interval widths converge to a constant.

```{r fig.width=8, fig.height=5, echo = FALSE, cache= TRUE}
ggplot(filter(pi_dat, distr == "Exponential"), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_line(aes(y = int_width_boot), color = "blue", size = 1.5) +
    ## geom_errorbar(aes(ymin=int_width - 2*int_width_se,
    ##                       ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("P.I Simulation, Exponential Distribution",
            subtitle = "Interval Widths. naive method (black), simulation method (blue)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    facet_wrap(~censoredp)
```

```{r fig.width=8, fig.height=5, echo = FALSE, cache= TRUE}
ggplot(filter(pi_dat, distr == "Loglogistic"), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_line(aes(y = int_width_boot), color = "blue", size = 1.5) +
    ## geom_errorbar(aes(ymin=int_width - 2*int_width_se,
    ##                   ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("P.I Simulation, Loglogistic Distribution",
            subtitle = "Interval Widths. naive method (black), simulation method (blue)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    facet_wrap(~censoredp)
```

```{r fig.width=8, fig.height=5, echo = FALSE, cache= TRUE}
ggplot(filter(pi_dat, distr == "Lognormal"), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_line(aes(y = int_width_boot), color = "blue", size = 1.5) +
    ## geom_errorbar(aes(ymin=int_width - 2*int_width_se,
    ##                   ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("P.I Simulation, Lognormal Distribution",
            subtitle = "Interval Widths. naive method (black), simulation method (blue)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    facet_wrap(~censoredp)
```

```{r fig.width=8, fig.height=5, echo = FALSE, cache= TRUE}
ggplot(filter(pi_dat, distr == "Weibull"), aes(x=sample_size, y=int_width)) +
    geom_point(size = 2) +
    geom_line(size = 1.5) +
    geom_line(aes(y = int_width_boot), color = "blue", size = 1.5) +
    ## geom_errorbar(aes(ymin=int_width - 2*int_width_se,
    ##                   ymax = int_width + 2*int_width_se), width=.1) +
    xlab("Sample Size (log scale)") +
    ylab("Interval Widths +/- 2 SEs") +
    ggtitle("P.I Simulation, Weibull Distribution",
            subtitle = "Interval Widths. naive method (black), simulation method (blue)") +
    scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) +
    facet_wrap(~censoredp)
```

<!-- ## Medians -->

<!-- ```{r, include = FALSE, cache= TRUE} -->
<!-- expq0 <- read.csv(paste0(path,"/","exponential_sim_quantile_ci_censoredp00.csv")) -->
<!-- expq1 <- read.csv(paste0(path,"/","exponential_sim_quantile_ci_censoredp03.csv")) -->
<!-- expq2 <- read.csv(paste0(path,"/","exponential_sim_quantile_ci_censoredp05.csv")) -->
<!-- expq_dat <- rbind(expq0, expq1, expq2) %>% -->
<!--     mutate(distr = "Exponential") -->
<!-- loglogq0 <- read.csv(paste0(path,"/", -->
<!--                             "loglogistic_sim_quantile_ci_scale025_censoredp00.csv")) -->
<!-- loglogq1 <- read.csv(paste0(path,"/", -->
<!--                             "loglogistic_sim_quantile_ci_scale025_censoredp03.csv")) -->
<!-- loglogq2 <- read.csv(paste0(path,"/", -->
<!--                             "loglogistic_sim_quantile_ci_scale025_censoredp05.csv")) -->
<!-- loglogq_dat <- rbind(loglogq0, loglogq1, loglogq2) %>% -->
<!--     mutate(distr = "Loglogisic") %>% -->
<!--     dplyr::select(-c(scale)) -->
<!-- lognormq0 <- read.csv(paste0(path,"/","lognormal_sim_quantile_ci_scale2_censoredp00.csv")) -->
<!-- lognormq1 <- read.csv(paste0(path,"/","lognormal_sim_quantile_ci_scale2_censoredp03.csv")) -->
<!-- lognormq2 <- read.csv(paste0(path,"/","lognormal_sim_quantile_ci_scale2_censoredp05.csv")) -->
<!-- lognormq_dat <- rbind(lognormq0, lognormq1, lognormq2) %>% -->
<!--     mutate(distr = "Lognormal") %>% -->
<!--     dplyr::select(-c(scale)) -->
<!-- weibullq0 <- read.csv(paste0(path,"/","weibull_sim_quantile_ci_scale2_censoredp00.csv")) -->
<!-- weibullq1 <- read.csv(paste0(path,"/","weibull_sim_quantile_ci_scale2_censoredp03.csv")) -->
<!-- weibullq2 <- read.csv(paste0(path,"/","weibull_sim_quantile_ci_scale2_censoredp05.csv")) -->
<!-- weibullq_dat <- rbind(weibullq0, weibullq1, weibullq2) %>% -->
<!--     mutate(distr = "Weibull") %>% -->
<!--     dplyr::select(-c(scale)) -->
<!-- quant_dat <- rbind(expq_dat, loglogq_dat, lognormq_dat, weibullq_dat) -->
<!-- ``` -->
<!-- ```{r fig.width=8, fig.height=8, echo = FALSE} -->
<!-- ggplot(filter(quant_dat), aes(x=sample_size, y=cov_prob)) + -->
<!--     geom_point(size = 2) + -->
<!--     geom_line(size = 1.5) + -->
<!--     geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se, -->
<!--                       ymax = cov_prob + 2*cov_prob_se), width=.1) + -->
<!--     xlab("Sample Size (log scale)") + -->
<!--     ylab("Coverage Probs +/- 2 SEs") + -->
<!--     ggtitle("C.I. for Median Simulation, by Distribution and Censoring", -->
<!--             subtitle = "Coverage Probabilities (zoomed in)") + -->
<!--     scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) + -->
<!--     geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) + -->
<!--     facet_grid(distr~censoredp) -->
<!-- ``` -->

<!-- ```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE} -->
<!-- ggplot(filter(quant_dat), aes(x=sample_size, y=cov_prob)) + -->
<!--     geom_point(size = 2) + -->
<!--     geom_line(size = 1.5) + -->
<!--     ylim(0,1) + -->
<!--     geom_errorbar(aes(ymin=cov_prob - 2*cov_prob_se, -->
<!--                       ymax = cov_prob + 2*cov_prob_se), width=.1) + -->
<!--     xlab("Sample Size (log scale)") + -->
<!--     ylab("Coverage Probs +/- 2 SEs") + -->
<!--     ggtitle("C.I. for Median Simulation, by Distribution and Censoring", -->
<!--             subtitle = "Coverage Probabilities (zoomed out)") + -->
<!--     scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) + -->
<!--     geom_hline(aes(yintercept = 0.90), colour="#BB0000", linetype="dashed", size = 1) + -->
<!--     facet_grid(distr~censoredp) -->
<!-- ``` -->

<!-- ```{r fig.width=8, fig.height=8, echo = FALSE, cache= TRUE} -->
<!-- ggplot(filter(quant_dat), aes(x=sample_size, y=int_width)) + -->
<!--     geom_point(size = 2) + -->
<!--     geom_line(size = 1.5) + -->
<!--     geom_errorbar(aes(ymin=int_width - 2*int_width_se, -->
<!--                       ymax = int_width + 2*int_width_se), width=.1) + -->
<!--     xlab("Sample Size (log scale)") + -->
<!--     ylab("Interval Widths +/- 2 SEs") + -->
<!--     ggtitle("C.I for Median Simulation, by Distribution and Censoring", -->
<!--             subtitle = "Interval Widths") + -->
<!--     scale_x_log10(breaks = c(20, 30, 50, 100, 250, 500, 1000, 2000)) + -->
<!--     geom_hline(aes(yintercept = 0.0), colour="blue", linetype="dashed", size = 1) + -->
<!--     facet_grid(distr~censoredp) -->
<!-- ``` -->


References:

Meeker, William Q., and Luis A. Escobar. Statistical methods for
reliability data. John Wiley & Sons, 2014. (Chapter 4, 8, and Appendix B)

Harrell, Frank E. Regression modeling strategies. Springer, 2015. (Chapter 17)



```{r }
sessionInfo()
```
