---
title: Introducing `ciTools`
author: John Haman and Matthew Avery
date: August 8, 2017
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing ciTools}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = F}
## setup 
set.seed(20170808)
knitr::opts_chunk$set(fig.width=8, fig.height=4) 
library(ciTools)
library(tidyverse)
library(knitr)
```

# Overview

The purpose of `ciTools` is to make it easier to do common
types of inference in R, particularly uncertainty bounds and
probability and quantile estimates. These are the tools researchers
use when comparing average system performance to requirements,
bounding future system performance, and estimating whether the system
will achieve specific thresholds that aren't necessarily average
performance.  Specifically, `ciTools` gives users access to one-line
commands that produce confidence intervals and prediction bounds for a
given design matrix. This matrix can be the observed data from a test
or a set of points that "span the space", allowing the analyst to
visualize system performance. For more information about spanning a
design space in **R**, see
[data_grid](https://www.rdocumentation.org/packages/modelr/versions/0.1.1/topics/data_grid)
or [crossing](http://tidyr.tidyverse.org/reference/expand.html).

`ciTools` makes these statistical quantities available through a
set of four functions that have a uniform syntax: `add_<*>(data,
model, ...)`. Users only need to learn one expression to get
started and can intuitively learn other functions when needed.

# Links
1. Another readme is available on our [GitHub
page](https://github.com/jthaman/ciTools).

1. Do you have a problem with `ciTools` or a feature that you would
like to see implemented? Create an Issue [on
the bug tracker](https://github.com/jthaman/ciTools/issues).

1. Comprehensive documentation (with examples!) is available
[here](https://www.rdocumentation.org/packages/ciTools/versions/0.1.0)

1. The main CRAN page for `ciTools` is
[here](https://cran.r-project.org/web/packages/ciTools/index.html)

# Using `ciTools`

We designed `ciTools` to make it easy and convenient to generate
intervals estimates when you're done building a model. Since the exact
formulation of a confidence interval depends on the type of
statistical model fit, figuring out how to construct a proper
confidence interval can be challenging. These functions automatically
identify the correct interval for you, provided your model is one of
the supported classes.

Here are the four main functions of `ciTools`
that you can use regardless of the type of model you made:

1. `add_ci(data, model, ...)` -- compute confidence intervals for
the fitted values of each row in `data` and append to `data`.
1. `add_pi(data, model, ...)` -- compute prediction intervals for
the fitted values of each row in `data` and append to `data`.
1. `add_probs(data, model, ...)` -- compute conditional response
probabilities for the fitted values of each row in `data` and append
to `data`.
1. `add_quantiles(data, model, ...)` -- compute
conditional response quantiles for the fitted values of each row in
`data` and append to `data`.

In each of the above functions, `model` is the model you've fit (of
class `lm`, `glm`, or `lmerMod`, which correspond to models fit
with the functions `lm`, `glm`, and `lmer` respectively), and
`data` is the matrix of data points for which you'd like
uncertainty estimates.

Each function returns `data` with your estimates appended to
facilitate plotting with `ggplot`. For those familiar with the
`modelr` package, they function the same way as
`add_predictions`. Another advantage is to make all of these
commands interoperable: they may be chained together to give all
the quantities of interest at once.

# Example

Here we will feature a common linear model in **R** that uses the
`cars` dataset. Uncertainty intervals in **R** for linear models
are well supported through the functions `predict.lm`, so the
functions we provide in `ciTools` are "wrappers" over `predict.lm`.


```{r }
my_data <- cars
glimpse(my_data)
```


A linear model that estimates stopping distance as a function of
speed:


```{r }
model <- lm(dist ~ speed, data = cars)
```

## Confidence Intervals

If we were interested in the average stopping distance as a
function of speed, we can generate a confidence interval using
`add_ci()`.  The output is another data frame with three new
columns: one for the model predictions, one for the lower
confidence bound, and one for the upper confidence bound:


```{r }
my_data_with_ci <- add_ci(my_data, model, names = c("lcb", "ucb"))
kable(head(my_data_with_ci, n =10), row.names = TRUE)
```


The data and the model fit can be inspected graphically with
`ggplot`:


```{r }
my_data_with_ci %>%
    ggplot(aes(x = speed, y = dist)) +
    geom_point(size = 2) +
    geom_line(aes(y = pred), size = 2, color = "maroon") +
    geom_ribbon(aes(ymin = lcb, ymax = ucb), fill =
    "royalblue1", alpha = 0.3) + 
    ggtitle("Stopping Distance vs. Car Speed: 95% Confidence Interval") +
    xlab("Car Speed (mph)") +
    ylab("Stopping Distance (ft)")
```

The red line is the model's estimated average stopping distance
across the different car speeds, and the blue region represents the
uncertainty in the average stopping distance. The default
confidence level is 95 percent. Users who desire a different
confidence level can use the option `alpha = *` to specify a custom
level. For example, if 80 percent intervals are desired, then
include `alpha = 0.2` in the `add_ci()` call.

## Prediction Intervals

Prediction intervals are similar to confidence intervals, but
instead of conveying the uncertainty in the estimated average,
prediction intervals convey the uncertainty in a new
observation. There is more uncertainty about a single new
observation than the average of all new observations, so prediction
intervals are wider than confidence intervals. To generate
prediction intervals, use `add_pi`:


```{r }
my_data_with_pi <- add_pi(my_data, model, names = c("lpb", "upb"))
```

The data frame is now larger because we have two new columns for
lower and upper prediction bounds tacked on to the end:


```{r }
kable(head(my_data_with_pi, n = 10), row.names = TRUE)
```



Here is what it looks like when we represent confidence intervals
and prediction intervals at the same time:

```{r }
my_data %>%
  add_ci(model, names = c("lcb", "ucb")) %>%
  add_pi(model, names = c("lpb", "upb")) %>%
    ggplot(aes(x = speed, y = dist)) +
    geom_point(size = 2) +
    geom_line(aes(y = pred), size = 2, color = "maroon") +
    geom_ribbon(aes(ymin = lpb, ymax = upb), fill = "orange2",
                alpha = 0.3) +
    geom_ribbon(aes(ymin = lcb, ymax = ucb), fill =
    "royalblue1", alpha = 0.3) + 
    ggtitle("Stopping Distance vs. Car Speed: 95% CI and 95% PI") +
    xlab("Car Speed (mph)") +
    ylab("Stopping Distance (ft)")
```

In the graph above, the blue confidence intervals show the
uncertainty in the model fit itself (maroon line), and the orange
prediction intervals shows where the model would predict 95% of new
responses (Stopping Distances) to fall.

## Response Probabilities

Often we want other quantities that depend on the conditional
predictive distribution. These include response-level probabilities
and response-level quantiles, which are accessed with the functions
`add_probs` and `add_quantile` respectively. 

For example, suppose in the `cars` data set, I want to know: For
each Speed what is the probability that a new Stopping Distance
will be less than 70 feet? This may be an important question if,
for example, you're in a car that is hurdling towards a cliff 70
feet away at some speed and you want to know what the probability
is that you will be able to stop the car before going over the
cliff. Luckily, with `add_probs()`, you can generate these
estimates quickly, allowing you to understand your chance of survival
before the problem becomes moot:


```{r }
my_data %>%
  add_probs(model, q = 70) %>%
    ggplot(aes(x = speed, y = prob_less_than70)) +
    geom_line(aes(y = prob_less_than70), size = 2, color = "maroon") +
    scale_y_continuous(limits = c(0,1)) +
    ggtitle("Probability Stopping Distance is Less Than 70") +
    xlab("Car Speed (mph)") +
    ylab("Pr(Dist < 70)")
```


The new argument `q = *` is used to specify the quantile (70 feet
in this case) used for computing the probabilities. It's clear that
the probability of surviving declines quickly after 15 mph. (This
data set is from the 1920s. It is safer to drive toward cliffs in
today's vehicles).

Another optional argument is `comparison`, which defaults to
`"<"`. In this example, we wanted to know the probability that we'd
be able to stop the car before it went over the cliff, which means
a stopping distance less than 70 feet. If we were to specify
`comparison = ">"`, then `add_probs()` would return the probability
that the stopping distance was greater than 70 feet.

## Response Quantiles

On the other hand, suppose my car is hurdling toward a cliff, and
I'm comfortable with stopping at whatever distance guarantees about
90% survivability given my speed. These distances are called
response quantiles, and what we wish to compute is the 0.9-quantile
(or the 90th percentile) of the distibution of Stopping Distances
*conditional* on my car's speed and the linear model. In `ciTools`
we use the function `add_quantile` with the argument `p = 0.9` to
calculate the 90th percentile of the predictive distibution for
each row in the data set. These quantiles will be parallel to the
bounds of the prediction intervals that we calculated previously.


```{r }
my_data %>%
  add_pi(model, names = c("lpb", "upb")) %>%
  add_quantile(model, p = 0.9) %>%
    ggplot(aes(x = speed, y = dist)) +
    geom_point(size = 2) +
    geom_line(aes(y = pred), size = 2, color = "maroon") +
    geom_line(aes(y = quantile0.9), size = 2, color = "forestgreen") + 
    geom_ribbon(aes(ymin = lpb, ymax = upb), fill = "orange2",
                alpha = 0.3) +
    ggtitle("Stopping Distance vs. Car Speed: 95% PI with 0.9-Quantile") +
    xlab("Car Speed (mph)") +
    ylab("Stopping Distance (ft)")
```

The 0.9 quantile lies slightly below the upper prediction bound,
which in this case is the same as the 0.975-quantile. The red line
is the prediction the linear model makes, which is the same as the
0.5-quantile in this case because our model assumes normally
distributed errors.

The examples above have taken each piece of analysis one step at a
time. However, because `ciTools` was built to be fully compatible
with the `tidyverse`, these functions can easily be chained or
"piped" together in clear, legible code:

```{r }
data_with_results <- my_data %>%
  add_ci(model) %>%
  add_pi(model) %>%
  add_probs(model, q= 70) %>%
  add_quantile(model, p = 0.9) 

kable(head(data_with_results))
```


# Scope of `ciTools`

`ciTools` handles more than just linear models but the workflow is
the same as the example above. Here is current status of
development:


| Models                   | Confidence Intervals | Prediction Intervals | Response Probabilities | Response Quantiles |
|--------------------------|:--------------------:|:--------------------:|:----------------------:|:------------------:|
| Linear                   |          [X]         |          [X]         |           [X]          |         [X]        |
| Log-Linear               |          [X]         |          [X]         |           [X]          |         [X]        |
| GLM                      |          [X]         |       [Poisson]      |        [Poisson]       |      [Poisson]     |
| Linear Mixed Model       |          [X]         |          [X]         |           [X]          |         [X]        |
| Log-Linear Mixed         |        [TODO]        |          [X]         |           [X]          |         [X]        |
| Survival                 |        [TODO]        |        [TODO]        |         [TODO]         |       [TODO]       |
| Generalized Linear Mixed |        [TODO]        |        [TODO]        |         [TODO]         |       [TODO]       |

* [X] = complete, available now

# Installation

Open up **R** and run:

1. `install.packages("ciTools")`
1. `library(ciTools)`
